---
title: "EDGE Buildings"
author: "Antoine Levesque, Pascal Führlich"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EDGE Buildings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
knitr::opts_knit$set(root.dir = rprojroot::find_package_root_file())
```


# Energy Demand GEnerator (EDGE)

## What is EDGE?
The Energy Demand GEnerator projects energy demand for buildings both at the
useful final energy level. It covers the global demand and five energy services--
space heating, space cooling,appliances and lighting (treated together) water heating
and cooking. In a nutshell, it computes the useful energy intensity of each service
in different regions, taking into account climate conditions as well as floor space
demand. It then projects energy demand based on the assumption that, as developing
countries grow economically, they will catch up the energy service demand of
developed countries. Further assumptions are made for the level beyond
the current level of economic development (saturation or lower growth).


## Documentation
The model documentation as well as the equation documentation can be found in
paper "How much energy will buildings consume in 2100? A global perspective within
a scenario framework" by Antoine Levesque, Robert C. Pietzcker, Lavinia Baumstark,
Simon De Stercke, Arnulf Grübler and Gunnar Luderer.

## How to use the model

### TLDR
```{r, eval = FALSE}
renv::restore() # install dependencies into ../renv-libraries
devtools::install(dependencies = FALSE) # install edgebuildings
targets::tar_make() # run all available calculations, generate all files
targets::tar_visnetwork() # get overview, show status, check if everything worked
```

### Setup
Install dependencies using [renv](https://rstudio.github.io/renv/):
```{r, eval = FALSE}
renv::restore()
```
The dependencies are installed in an isolated package library that is used exclusively for EDGE Buildings. The path to
this package library can be set in the `.Renviron` file, it defaults to `../renv-libraries`.

### Running
EDGE Buildings is using [targets](https://docs.ropensci.org/targets/) for running calculations and caching. Note that
`targets` functions must be run in this project's root directory. The available calculations/targets are defined in
`_targets.R`. Here's an interactive visualization. You can zoom in to read the labels and drag everything around.
```{r, eval = FALSE}
targets::tar_visnetwork()
```
Each node in this graph is called a `target` and represents a function, a calculation or a file. In order to run a
calculation or generate a file use `targets::tar_make(<target>)`. For example, to generate all files that are used as
input in REMIND:
```{r, eval = FALSE}
targets::tar_make("remindInputData")
```
To run all available and outdated calculations run `targets::tar_make()`. The `targets` package will automatically mark
targets as "outdated" if an upstream target becomes outdated or if an underlying function or file changes. With
`targets::tar_make` only outdated targets are re-calculated, up-to-date targets are instead loaded from cache.

### Structure of the model
The file computing the energy intensities as well as making the projections is
scripts/buildings_projections.R. This script calls the getData function to load
the necessary data for the computation. Most of the functions in the functions folder
prepare intermediary data for the projections. They may as well call functions preparing
intermediary data from the raw data.
For instance, the buildings_projections.R script call "FloorspaceBuild", the floorspace
projections for residential, commercial and buildings floorspace demand. "FloorspaceBuild",
in turn, calls "FloorspaceResidential" which makes the residential floorspace projections,
and "ShareFloorCommercial" which determines the share of commercial area. Residential 
floorspace projections are based on floorspace data ("Floorspace" <- "Daioglou", "FloorspaceEURIND").

The data_internal folder stores the scenarios assumptions, the plot aesthetics as well
as the mappings.

The external data consists of two parts: first, the data that is called for via the madrat libraries.
These libraries require the data to be organised in a specific way 
coherent with their functions. Second, the data that is not called via these libraries and whose
path is called explicitly in the EDGE code. Both folders are not part of the EDGE model.

### Configuration
In the config() function, set the madrat data folder path by changing
the argument mfolder. 
In the config() function, set the madrat regional mapping path (from mfolder)
by changing the argument rmapping. 
In the config() function, set the external data folder path by changing
the argument external_data. 

Start the main.R file to load all the libraries and options necessary for the model
to run correctly.

### getData
getData is a central function of the model as all data processing functions are called
through it. The name of all data processing functions takes the form getNNN(), with NNN,
the specific name of the function. Once they are called through getData("NNN"), a cache file
is produced and stored in the cache folder (created by main.R if not existent). This cache
allows reducing the model time run by avoiding the whole function to be run on a second call.
Simply the cache file is called. If the user wishes to run the function as a default when
getData is called, the getData argument fromCache should be set to FALSE. After a function
has been changed, and if fromCache is set to TRUE, the cache file should be deleted for the
data processing function to be called from getData(NNN).

## License
GNU AGPLv3, see LICENSE.md

## How to contribute
We are interested in working with you! Just contact us through
[GitHub](https://github.com/pik-piam/edgeBuildings.git) or by mail
(robin.hasse@pik-potsdam.de) if you have found and/or fixed a bug, developed a new
model feature, have ideas for further model development, suggestions for
improvements or anything else. We are open to any kind of contribution. Our aim
is to develop an open, transparent and meaningful model. Join us doing so!
